---
title: "EPC data:  \nTechnical documentation"
output: 
  word_document:
    reference_doc: SERL_word_template_portrait1.docx
    toc: no
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(data.table)
library(lubridate)
library(knitr)
library(captioner)
library(flextable)
library(officer)

# Import all filenames, locations, source function files
source("D:/Users/ucldata/observatory_release_pre_processing/observatoryData/scripts/edition02/setup_edition02.R")

```

```{r defineInputs, include=FALSE}

load_epc <- TRUE # FALSE if using previously saved .RData
write_epc <- TRUE # TRUE if we want to save a new csv data file

```

```{r functions, include = FALSE}

get.cases <- function(i, d = epc) {
  v <- colnames(d)[i]
  tmp <- epc[, .N, keyby = v]
  tmp2 <- data.table(Variable = rep(v, nrow(tmp)),
                     Value = tmp[, 1],
                     Cases = tmp[, 2])
  colnames(tmp2) <- c("Variable", "Value", "Cases")
  return(tmp2)
}

my.flex <- function(t, autofit = FALSE) {
  ft <- flextable(t)
  ft <- theme_booktabs(ft)
  ft <- theme_zebra(ft, 
                    odd_header = rgb(84/255, 141/255, 212/255, 1),
                    even_body = "transparent",
                    odd_body = rgb(242/255, 242/255, 242/255, 1))
  ft <- color(ft, color = "white", part = "header")
  ft <- italic(ft, italic = TRUE, part = "header")
  ft <- bold(ft, bold = FALSE, part = "header")
  if(autofit == TRUE) {
    ft <- autofit(ft)
  } else{
    ft <- set_table_properties(ft, layout = "autofit")
  }
  return(ft)
}

survey.flex <- function(t, autofit = TRUE) {
  ft <- flextable(t)
  ft <- theme_booktabs(ft)
  ft <- border(ft, border.top = fp_border(color = "gray") )
  ft <- bg(ft, bg = rgb(84/255, 141/255, 212/255, 1), 
            part = "header")
  ft <- color(ft, color = "white", part = "header")
  ft <- italic(ft, italic = TRUE, part = "header")
  ft <- bold(ft, bold = FALSE, part = "header")

  ft <- set_table_properties(ft, width = 1, layout = "autofit")
  return(ft)
}

```

```{r loadData, include=FALSE}

if(load_epc == TRUE) {
  epc_orig <- fread(epc_file)
  epc <- copy(epc_orig)
  setnames(epc, old = "puprn", new = "PUPRN")
  save(epc,
       file = paste(location_processed, 
                    epc_output_name, ".RData", sep = ""))
} else {
  load(paste(location_processed, 
             epc_output_name, ".RData", sep = ""))
}

if(write_epc == TRUE) {
  fwrite(epc, 
         file = paste(location_processed, epc_output_name, ".csv", sep = ""))
}

```

```{r}
info_tab
```


```{r captions, include = FALSE}

tab_caps(name = "epc_fields", 
         caption = "All EPC variables, the number of unique values found for each variable, the variable (R) class, and an example from the dataset. In some cases examples are invented (for statistical disclosure control) as the count for unique reads is very low; in these instances examples take the same format as the real data.")

tab_caps(name = "summary_short", 
         caption = "The number and percent of each value found in the dataset for each variable with fewer than 10 unique values found. Note that number is rounded down to the nearest 10 for statistical disclosure control, and percent is the rounded number as a percent of total (so may not sum to 100%). ")

```

```{r summary, include = FALSE}

ncol_epc <- ncol(epc)
nrow_epc <- nrow(epc)

# Create a summary of the values for each variable

test_list <- lapply(2:ncol_epc, get.cases)

summary <- rbindlist(test_list)


summary[, Percent := round(Cases / nrow_epc * 100, 2)]


epc_fields <- data.table(Variable = colnames(epc),
                         class = lapply(epc, class),
                         example = rep(NA_character_, ncol_epc),
                         position = 1:ncol(epc)
                         )
setkey(epc_fields, Variable)
tmp <- summary[, .N, keyby = Variable]
epc_fields <- tmp[epc_fields]
setnames(epc_fields, 
         old = "N", 
         new = "uniqueValues")

epc_fields[Variable == "PUPRN", 
           uniqueValues := nrow_epc]

setkey(epc_fields, position)

set.seed(3)
for(x in 1:ncol_epc) {
  epc_fields[x, example := epc[round(runif(n = 1, min = 1, max = nrow_epc), 0), 
                               x, 
                               with = FALSE]]
}

# Replace cases that gave missing, NA, etc. values
epc_fields[Variable == "floorLevel", 
           example := "Ground"]
epc_fields[Variable == "flatTopStorey", 
           example := "Y"]
epc_fields[Variable == "flatStoreyCount", 
           example := "2"]
epc_fields[Variable == "numberOpenFireplaces", 
           example := "0"]
epc_fields[Variable == "floorEnergyEff", 
           example := "Very Good"]
epc_fields[Variable == "floorEnvEff", 
           example := "Good"]
epc_fields[Variable == "sheatingEnergyEff", 
           example := "N/A"] # No other responses
epc_fields[Variable == "sheatingEnvEff", 
           example := "N/A"] # No other responses
epc_fields[Variable == "roofEnergyEff", 
           example := "Good"]
epc_fields[Variable == "mainFuel", 
           example := "oil (not community)"]
epc_fields[Variable == "heatLossCorridoor", 
           example := "unheated corridor"]
epc_fields[Variable == "unheatedCorridorLength", 
           example := "4.998"]
epc_fields[Variable == "floorHeight", 
           example := "2.40"]

epc_fields[Variable == "glazedArea", 
           example := "Much More Than Typical"]
epc_fields[Variable == "lowEnergyLighting", 
           example := "67"]
epc_fields[Variable == "photoSupply", 
           example := "40"]
epc_fields[Variable == "lowEnergyFixedLightCount", 
           example := "4"]

# Check not revealing examples with too low a count

for(i in 1:nrow(epc_fields)) {
  epc_fields[i, N := epc[get(epc_fields[i, Variable]) == epc_fields[i, example], .N]]
}

epc_fields[N < 10, ]

epc_fields[Variable == "PUPRN", 
           example := "1ABC2DE3"] # invented, check not in the data

tmp <- epc[, .N, keyby = inspectionDate]
tmp[, inspectionDate := as.Date(inspectionDate, format = "%d/%m/%Y")]
setkey(tmp, inspectionDate)

epc_fields[Variable == "inspectionDate", 
           example := "13/05/2008"] # invented, check not in the data

tmp <- epc[, .N, keyby = constituency]
setkey(tmp, N)

epc_fields[Variable == "constituency", 
           example := "E14001013"] # Count > 10

tmp <- epc[, .N, keyby = lodgementDate]
tmp[, lodgementDate := as.Date(lodgementDate, format = "%d/%m/%Y")]
setkey(tmp, lodgementDate)

epc_fields[Variable == "lodgementDate", 
           example := "01/04/2009"] # invented, check not in the data


tmp <- epc[, .N, keyby = energyConsumptionCurrent]
setkey(tmp, N)

epc_fields[Variable == "energyConsumptionCurrent", 
           example := "234"] # Count > 10

tmp <- epc[, .N, keyby = heatingCostCurrent]
setkey(tmp, N)
setkey(tmp, heatingCostCurrent)

epc_fields[Variable == "heatingCostCurrent", 
           example := "270"] # invented, check not in the data


tmp <- epc[, .N, keyby = heatingCostPotential]
setkey(tmp, N)

epc_fields[Variable == "heatingCostPotential", 
           example := "356"] # Count > 10

tmp <- epc[, .N, keyby = hotWaterCostPotential]
setkey(tmp, N)

epc_fields[Variable == "hotWaterCostPotential", 
           example := "57"] # Count > 10

tmp <- epc[, .N, keyby = totalFloorArea]
setkey(tmp, N)

epc_fields[Variable == "totalFloorArea", 
           example := "80.00"] # Count > 10

tmp <- epc[, .N, keyby = unheatedCorridorLength]
setkey(tmp, N)
setkey(tmp, unheatedCorridorLength)

epc_fields[Variable == "unheatedCorridorLength", 
           example := "5.189"] # invented, check not in the data

tmp <- epc[, .N, keyby = photoSupply]
setkey(tmp, N)

epc_fields[Variable == "photoSupply", 
           example := "0"] # Count > 10

epc_fields[Variable == "lodgementDatetime", 
           example := "02/10/2010 14:51"]

# Check fixed
for(i in 1:nrow(epc_fields)) {
  epc_fields[i, N := epc[get(epc_fields[i, Variable]) == epc_fields[i, example], .N]]
}

epc_fields[N < 10, ]


for_short_summary <- epc_fields[uniqueValues < 10, Variable]

summary_short <- summary[Variable %in% for_short_summary | Variable == "PUPRN",]

# Round cases DOWN to nearest 10 for SDC
summary_short[, Cases := 10 * round((Cases - 4.999) / 10)]
summary_short[, Percent := round(Cases / epc_fields[Variable == "PUPRN", uniqueValues] * 100, 1)]

# Tables

## epc_fields
epc_fields[, `:=`(position = NULL,
                  N = NULL)]
colnames(epc_fields) <- c("variable", "n unique values", "class", "example")
epc_fields <- survey.flex(epc_fields)


## summary_short
colnames(summary_short) <- c("variable", "value", "number (rounded)", "percent")
summary_short <- survey.flex(summary_short)
summary_short <- merge_v(summary_short, j = ~ variable)
summary_short <- fix_border_issues(summary_short)


```

# Introduction

This document describes the England and Wales Energy Performance Certificate (EPC) data collected for SERL participants, stored in the file *`r paste(get.serl.filename("epc_data", release_version), ".csv", sep = "")`*. The data contains `r ncol_epc` columns and `r nrow_epc` rows (one row per participant with available EPC data). This document lists the EPC variables available along with basic information about the values for each variable such as number of unique values and statistics for numerical variables. A guide to the variables is available [here](https://epc.opendatacommunities.org/docs/guidance#glossary_domestic).

Data were collected with the Domestic Energy Performance Certificates API using the house number and postcode (details [here](https://epc.opendatacommunities.org/docs/api/domestic)). Where more than one EPC is registered for an address the most recent is provided.  

The data have not been modified from the original source except for the removal of address data (replaced with our PUPRN (a unique identifier) used in the other datasets). Note that data quality analysis has not been performed for this dataset.

# Data summary

Table 1 lists all variables currently available in the SERL EPC dataset. The number of unique values is given, alongside the R data class and an example value from the dataset. 

For variables with fewer than 10 unique values in the EPC dataset, Table 2 shows the number of records with each value and the percent with this value (or non-value in the case of N/A or 'NO DATA!' etc.). We also include PUPRN to show the number of records. 


`r tab_caps("epc_fields")`
```{r}
epc_fields
```



`r tab_caps("summary_short")`
```{r}
summary_short
```





```{r save, include = FALSE}

save(nrow_epc,
     ncol_epc,
     file = paste(epc_stats_file, sep = "")
)

```












