---
title: 'Climate data: technical documentation'
output: 
  word_document:
    reference_doc: SERL_word_template_portrait1.docx
    toc: no
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(data.table)
library(lubridate)
library(knitr)
library(captioner)

# Import all filenames, locations, source function files
source("N:/R/observatoryData/scripts/setup_v2020_08.R")

```

```{r defineInputs, include = FALSE}

load_climate <- TRUE
write_climate <- FALSE

```

```{r loadData, include=FALSE}

if(load_climate == TRUE) {
  climate_orig_list <- lapply(1:length(climate_file_dates), 
                       function(x) {
                         fread(paste(climate_location, 
                                     "Climate Data ",
                                     climate_file_dates[x],
                                     ".csv", 
                                     sep = "")) 
                       }
)

climate_orig <- rbindlist(climate_orig_list)

climate <- copy(climate_orig)

climate[, analysis_date := ymd(analysis_date)]
climate[, date_time := ymd_hms(date_time)]


setnames(climate, 
         old = c("grid_cell",
                 "analysis_date",
                 "date_time"),
         new = c("gridCell",
                 "analysisDate",
                 "utcDateTime")
         )

save(climate,
     file = paste(climate_location, climate_output_name, ".RData", sep = "")
     )

} else {
  load(paste(climate_location, climate_output_name, ".RData", sep = ""))
}

if(write_climate == TRUE) {
  fwrite(climate,
         file = paste(location_processed, climate_output_name, ".csv", sep = ""))
}


```

```{r}
info.tab
```

```{r captions, include = FALSE}

tab_caps(name = "climate_fields", 
         caption = "Climate data fields.")

tab_caps(name = "stats", 
         caption = "Data statistics")

```


# Introduction

This document describes the climate data provided with July 2020 SERL smart meter data release which originates from the Copernicus/ECMWF [ERA5 hourly reanalysis data](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-single-levels?tab=overview). 

More variables will be provided in future releases, but at the moment only temperature at 2 metres is part of the data. It can be linked with SERL participants using the 'gridCell' variable. The filename of this dataset is *`r paste(climate_output_name, ".csv", sep = "")`*. This documentation will be expanded in future as more climate variables are provided with SERL datasets. 

# Overview

The climate data has `r ncol(climate)` columns and `r nrow(climate)` rows. This is reanalysis data based on recorded data from many weather stations across GB, at a horizontal resolution of 0.25 x 0.25 degrees latitute and longitude (approximately 28 sq. km). More information about spacial resolution is available [here](https://confluence.ecmwf.int/display/CKB/ERA5%3A+What+is+the+spatial+reference). ERA5 documentation is available [here](https://confluence.ecmwf.int/display/CKB/ERA5%3A+data+documentation).

The fields provided in the current SERL climate dataset are described in Table 1. 

```{r fieldsTable, include = FALSE}

nrow_climate <- nrow(climate)
ncol_climate <- ncol(climate)

climate_fields <- data.table(columnNames = colnames(climate),
                            description = rep(NA_character_, ncol_climate),
                            units = rep(NA_character_, ncol_climate),
                            class = lapply(climate, class),
                            example = rep(NA_character_, ncol_climate))


climate_fields[columnNames == "gridCell",
               `:=`(description = "Grid cell for linking to participant data",
                    units = "NA",
                    example = "38_31")]

climate_fields[columnNames == "analysisDate",
               `:=`(description = "Date the data refers to",
                    units = "NA",
                    example = "2018-08-01")]

climate_fields[columnNames == "utcDateTime",
               `:=`(description = "Time and date of the data in UTC",
                    units = "%Y-%m-%d %H:%M:%S",
                    example = "2018-08-01 00:00:00")]


climate_fields[columnNames == "temperature_2_metres_K",
               `:=`(description = "Temperature at 2 metres from the surface",
                    units = "deg K",
                    example = "285.12")]


climate_fields[columnNames == "temperature_2_metres_C",
               `:=`(description = "Temperature at 2 metres from the surface converted into Celsius",
                    units = "deg C",
                    example = "12.12")]

colnames(climate_fields) <- c("Field", "Description", "Units", "Class", "Example value")

```

`r tab_caps("climate_fields")`
```{r}
kable(climate_fields)
```

Table 2 provides some basic information and statistics about the data.

```{r stats, include = FALSE}

stats <- data.table(Statistic = c("Number of grid cells",
                                  "Earliest read date",
                                  "Latest read date",
                                  "Lowest temperature (deg C)",
                                  "Highest temperature (deg C)",
                                  "Mean temperature (deg C)"),
                    Value = rep(NA_character_, 6)
)

stats[Statistic == "Earliest read date", 
      Value := as.character(climate[1, analysisDate])]
stats[Statistic == "Latest read date", 
      Value := as.character(climate[nrow(climate), analysisDate])]
stats[Statistic == "Lowest temperature (deg C)", 
      Value := climate[, min(temperature_2_metres_C)]]
stats[Statistic == "Highest temperature (deg C)", 
      Value := climate[, max(temperature_2_metres_C)]]
stats[Statistic == "Mean temperature (deg C)", 
      Value := climate[, round(mean(temperature_2_metres_C), 2)]]
stats[Statistic == "Number of grid cells", 
      Value := length(unique(climate$gridCell))]

```

`r tab_caps("stats")`
```{r}
kable(stats)
```

```{r save, include = FALSE}

save(nrow_climate,
     ncol_climate,
     file = climate_stats_file
)

```

